### 2019-03-09 manuscript/introduction.tex
- restoring introduction after bad commit
read through Results

%B I got a strange conflict here (first few paragraphs) with a seemingly long-ago committed repo version of introduction? The excised text is pasted at the bottom, and commented out, just in case.

@@

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility causes plants to reject their own pollen, and its loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}. % Gordon: Really? I find this hard to believe... %B:  sigh.
Previous analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that unobserved states drive diversification patterns \citep{freyman_2018}.
Despite their apparent macroevolutionary role in diversification, breeding systems are clearly not the sole trait determining diversification rates.
Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
%B: I did not find this confusing, but it worries me that Gordon did. Consequently, I've ejected this sentence, now focusing on SI/SC only.
%The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
Moreover, polyploidy is not only a mere correlate of breeding system, but may directly cause the loss of self-incompatibility (SI) \citep{stout1942, lewis1947}, and facilitate the transition to selfing.

Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
This precludes the existence of SI polyploids in the most widespread system of SI, and creates a correlation between polyploidy and SC.
% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
%B: not confused here; some SI is broken down by ploidy. Therefore, there is a correlation with coefficient between 0 and 1, and not 1.0 
% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems. I don't think the de Vos paper is dealing with SI systems. If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying. 
% R- Boris what do you think is the best way to proceed here? Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
%B: The apparent problem for Gordon is that devos2014 estimates *SSE parameters for heterostyly and homostyly, and not homomorphic SI. That's not a concern, I think, they are all "SI systems." And Dioecy etc. is not classically considered a "breeding system", but a "sexual system". That may be a part of the confusion. It's really not Gordon's problem, and we could do something to yet again re-define these terms? Emma?

Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
First, it may be of particular interest to establish their joint influence on lineage diversification.
It is possible that combinations of breeding system and ploidy do not result in predictable additive effects on speciation and extinction rates.
Second, the two traits may affect each other's state transitions, including their magnitude and ordering of transitions.
For example, losses of SI in diploids could be commonly caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without in increase in ploidy, and then a subsequent diploidization. 
%Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids
\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the one-step pathway over long timescales, but generally proceeds in one-step via polyploidization of SI species over short timescales.
A vast accumulation of data and methods that allow for diversification rate differences now allow far more powerful approach to such problems. % E: Technically, we allow for div effects in the rate estimates, but not in the pathways.  If SC-D have so much extinction that the step to SC-P never happens, the old pathway methods won't see that.  I'll think about if this is easy to fix.
% And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} % E: This is a great question.  We don't answer it with this round of analyses.  Maybe when we have more time for the revision?  <- FIXME?

@@

%>>>>>>> aa786e0237f5a090b69a64b415f317a49723f88c
%Species accumulate around the tree of life at different rates.
%A possible explanation for this phenomenon is that species possess traits that cause unequal rates of diversification. %B: new.
%The search for influential traits has been accelerated by a dramatic increases in data and an advance in methods, which now make use of far more information from phylogenetic trees \citep{maddison_2007}. 
%Despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are enabling the discovery of traits that alter the diversification patterns.
%Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
%This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process \citep[see][]{caetano_2018, herrera_2018}.
%Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
%Two exceedingly well-studied traits, ploidy and breeding system, are of particular interest. 
%Their individual state-dependent contributions to the process of diversification, as well as influence on each others' trait transitions, have long been the focus of speculation and study \citep{stebbins1950}.
%Whole genome duplication is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
%Such changes is ploidy have the potential to affect many other phenotypes, and therefore alter a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
%The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}. 
%Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citep{mayrose_2011, mayrose_2015}. 
%A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
%\citet{landis_2018}, for example, found 106 whole genome duplications within angiosperms that, combined with a state-dependent diversification model, resulted in approximately 60\% of whole genome duplication events potentially enhancing net diversification.
%The mere occurrence of such ancestral whole genome duplications points to an possibly important role of genome downsizing and diploidization in the diversification process \citep{soltis_2015diploidization, dodsworth_2015}. 
%The presence of a significant and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized. 
%Under this scenario, rather than being a macroevolutionary dead-end, polyploidy may instead be positively associated with the diversification process.

%Ploidy has long been thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
%The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
%In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility (SC) \citep{stout1942, lewis1947}.

>>>>

%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.

@@

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility causes plants to reject their own pollen, and its loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}. % Gordon: Really? I find this hard to believe... %B:  sigh.
Previous analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that unobserved states drive diversification patterns \citep{freyman_2018}.
Despite their apparent macroevolutionary role in diversification, breeding systems are clearly not the sole trait determining diversification rates.
Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
%B: I did not find this confusing, but it worries me that Gordon did. Consequently, I've ejected this sentence, now focusing on SI/SC only.
%The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
Moreover, polyploidy is not only a mere correlate of breeding system, but may directly cause the loss of self-incompatibility (SI) \citep{stout1942, lewis1947}, and facilitate the transition to selfing.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
This precludes the existence of SI polyploids in the most widespread system of SI, and creates a correlation between polyploidy and SC.
% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
%B: not confused here; some SI is broken down by ploidy. Therefore, there is a correlation with coefficient between 0 and 1, and not 1.0 
% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems. I don't think the de Vos paper is dealing with SI systems. If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying. 
% R- Boris what do you think is the best way to proceed here? Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
%B: The apparent problem for Gordon is that devos2014 estimates *SSE parameters for heterostyly and homostyly, and not homomorphic SI. That's not a concern, I think, they are all "SI systems." And Dioecy etc. is not classically considered a "breeding system", but a "sexual system". That may be a part of the confusion. It's really not Gordon's problem, and we could do something to yet again re-define these terms? Emma?

Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
First, it may be of particular interest to establish their joint influence on lineage diversification.
It is possible that combinations of breeding system and ploidy do not result in predictable additive effects on speciation and extinction rates.
Second, the two traits may affect each other's state transitions, including their magnitude and ordering of transitions.
For example, losses of SI in diploids could be commonly caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without in increase in ploidy, and then a subsequent diploidization. 
%Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids
\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
A vast accumulation of data and methods that allow for diversification rate differences now allow far more powerful approach to such problems.

@@
